%Acknowledgments

\renewcommand{\baselinestretch}{2}
\small\normalsize
\hbox{\ }
 
\vspace{-.65in}

\begin{center}
\large{Acknowledgments} 
\end{center} 

\vspace{1ex}

I owe my deepest gratitude to everyone who has supported me and provided me help during my graduate school experience. 

First, I would like to thank my academic advisor, Professor Rajeev Barua, for giving me the opportunity to work in a challenging and rigorous academic environment during my master's program. Having worked in his compiler research group as an undergraduate during the summer before I started my master's, I knew that continuing my research with him was the right decision. His invaluable advice and feedback has allowed me to learn and accomplish more than I could have imagined during the past year and a half. It has been a pleasure to work with such an extraordinary individual.

I would also like to thank Michael Ferguson, our program manager at the Laboratory for Telecommunication Sciences (LTS) in College Park, Maryland, for helping me complete the Chapel implementation portion of this project. I am deeply grateful that Michael was able to provide a place for me to work at LTS when possible and even offer to run tests and collect data for me remotely. His expertise in the Chapel programming language has really helped me understand how to turn our theory into practice. 

I would like to thank Darren Smith, who developed the theory and preliminary implementation for this project before he graduated and I took over. During the summer of 2013, Darren spent time to bring me up to speed on his work in order for me to understand his reasoning and to ensure a smooth transition. I would like to thank Joshua Koehler, who helped me implement all of our Chapel benchmarks that we used to test our optimization. Joshua worked extremely hard and was always available to talk about new ideas and directions that we could take this research. 

I would like to also thank both Professors Manoj Franklin and Charalampos Papamanthou for serving on my thesis advisory committee and spending their valuable time reviewing my manuscript.

Finally, I would like to thank my family -- my mother, father, and sister -- who have given me strength and support throughout my entire life, as well as my friends who always knew how to bring positivity into my life when school was tough. 

\begin{comment}
I owe my gratitude to all the people who have made this thesis possible and because of whom my graduate experience has been one that I will cherish forever.

First and foremost I'd like to thank my advisor, Professor Rajarshi Roy for giving me an invaluable opportunity to work on challenging and extremely interesting projects over the past four years. He has always made himself available for help and advice and there has never been an occasion when I've knocked on his door and he hasn't given me time. It has been a pleasure to work with and learn from such an extraordinary individual.

I would also like to thank my co-advisor, Dr. Parvez Guzdar. Without his extraordinary theoretical ideas and computational expertise, this thesis would have been a distant dream. Thanks are due to Professor Robert Gammon, Professor Edward Ott and Professor Thomas Antonsen for agreeing to serve on my thesis committee and for sparing their invaluable time reviewing the manuscript.

My colleagues at the nonlinear optics laboratory have enriched my graduate life in many ways and deserve a special mention. David DeShazer helped me start-off by rewriting the basic simulation code in a user-friendly format. Christian Silva provided help by setting up the GRENOUILLE apparatus and performing some of the simulations. My interaction with  Rohit Tripathi, Ryan McAllister, Vasily Dronov, Min-Young Kim, Elizabeth Rogers, William Ray, Jordi Garcia Ojalvo, Riccardo Meucci, Atsushi Uchida, and Fabian Rogister has been very fruitful. I'd also like to thank Wing-Shun Lam and Benjamin Zeff for providing the LaTex style files for writing this thesis.

I would also like to acknowledge help and support from some of the staff members. Donald Martin's technical help is highly appreciated, as is the computer hardware support from Edward Condon, LaTex and software help from Dorothea Brosius and purchasing help from Nancy Boone.

I owe my deepest thanks to my family - my mother and father who have always stood by me and guided me through my career, and have pulled me through against impossible odds at times. Words cannot express the gratitude I owe them. I would also like to thank Dr. Mohan Advani, Dr. Vasudeo Paralikar and Dr. Vinod Chaugule who are like family members to me.

My housemates at my place of residence have been a crucial factor in my finishing smoothly. I'd like to express my gratitude to Sivasankar Pandeti, Jayakumar Patil, Amit Trehan and Punyaslok Purakayastha for their friendship and support.

I would like to acknowledge financial support from the Office of Naval Research (ONR), Physics, for all the projects discussed herein.

It is impossible to remember all, and I apologize to those I've inadvertently left out.

Lastly, thank you all and thank God!
\end{comment}
